---
title: "Hayman Fire Recovery"
author: "Madeline Bean"
date: "10/3/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, warning=F,message=F}
library(tidyverse)
library(tidyr)
library(ggthemes)
library(lubridate)
library(ggpubr)

# Now that we have learned how to munge (manipulate) data
# and plot it, we will work on using these skills in new ways

knitr::opts_knit$set(root.dir='..')
```


```{r dataread, warning=F,message=F}
####-----Reading in Data and Stacking it ----- ####
#Reading in files
files <- list.files('data',full.names=T)


#Read in individual data files
ndmi <- read_csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')


ndsi <- read_csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read_csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>%
  gather(key='site',value='value',-DateTime,-data) %>%
  filter(!is.na(value))


```




## Question 1) 

What is the correlation between NDVI and NDMI? - here I want you to
convert the full_long dataset in to a wide dataset using the 
function "spread" and then make a plot that shows the correlation s a
function of if the site was burned or not (x axis should be ndmi)
You should exclude winter months and focus on summer months

```{r}
#changing data into wide
full_wide <- spread(data=full_long,key='data',value='value') %>%
  filter_if(is.numeric,all_vars(!is.na(.))) %>%
  mutate(month = month(DateTime),
         year = year(DateTime))

#filtering summer months
summer_only <- filter(full_wide,month %in% c(6,7,8,9))

#plotting summer months of variables
ggplot(summer_only,aes(x=ndmi,y=ndvi,color=site)) + 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.8))

```

There is a strong, positive correlation between NDMI and NDVI. 

## Question 2 

2) What is the correlation between average NDSI (normalized
 snow index) for January - April and average NDVI for June-August?
In other words, does the previous year's snow cover influence vegetation
 growth for the following summer?
 
```{r warning=F, message=F}
#variable ndsi months
var.snow_months <- c(1,2,3,4)

#variable ndvi months
var.growth_months <- c(6,7,8)

#mean NDSI per year
ndsi_avg <- full_wide[c("DateTime", "ndsi", "month", "year", "site")] %>%
  filter(month %in% var.snow_months) %>%
  group_by(site, year) %>%
  summarize(ndsi_avg=mean(ndsi))

#mean NDVI per year
ndvi_avg <- full_wide[c("DateTime", "ndvi", "month", "year", "site")] %>%
  filter(month %in% var.growth_months) %>%
  group_by(site, year) %>%
  summarize(ndvi_avg=mean(ndvi))

#combining NDVI and NDSI into one dataset
combined <- inner_join(ndvi_avg, ndsi_avg)

#correlation
cor(combined$ndvi_avg, combined$ndsi_avg)

ggplot(combined, aes(x=ndsi_avg, y=ndvi_avg, color=site)) +
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.8))
  
```
 
There is a low positive correlation of 0.180 between snow coverage and vegetation. This means that the previous year's snow cover has little effect on the vegetation growth for the following summer and that this correlation could be from other factors in the environment. Looking at the graph, you can see a difference in correlation between NDVI and NDSI burned and un-burned sites. 


## Q3

How is the snow effect from question 2 different between pre- and post-burn
and burned and unburned? 

```{r warning=F, message=F}
#snow effect pre-burn
preburn <- c(1984:2001)

#snow effect post-burn
postburn <- c(2003:2019)

#snow effect average for preburn
ndsi_preburn_avg <- full_wide[c("DateTime", "ndsi", "month", "year", "site")] %>%
  filter(month %in% var.snow_months) %>%
  filter(year %in% preburn) %>%
  group_by(site, year) %>%
  summarize(ndsi_preburn_avg=mean(ndsi))

#vegetation effect average for preburn
ndvi_preburn_avg <- full_wide[c("DateTime", "ndvi", "month", "year", "site")] %>%
  filter(month %in% var.growth_months) %>%
  filter(year %in% preburn) %>%
  group_by(site, year) %>%
  summarize(ndvi_preburn_avg=mean(ndvi))

#combing preburn ndsi and ndvi into one data set
combine_preburn <- inner_join(ndsi_preburn_avg, ndvi_preburn_avg)

#correlation for preburn
cor(combine_preburn$ndsi_preburn_avg, combine_preburn$ndvi_preburn_avg)

#snow effect average for postburn
ndsi_postburn_avg <- full_wide[c("DateTime", "ndsi", "month", "year", "site")] %>%
  filter(month %in% var.snow_months) %>%
  filter(year %in% postburn) %>%
  group_by(site, year) %>%
  summarize(ndsi_postburn_avg=mean(ndsi))

#vegetation effect average for postburn
ndvi_postburn_avg <- full_wide[c("DateTime", "ndvi", "month", "year", "site")] %>%
  filter(month %in% var.growth_months) %>%
  filter(year %in% postburn) %>%
  group_by(site, year) %>%
  summarize(ndvi_postburn_avg=mean(ndvi))

#combing postburn ndsi and ndvi into one data set
combine_postburn <- inner_join(ndsi_postburn_avg, ndvi_postburn_avg)

#correlation for postburn
cor(combine_postburn$ndsi_postburn_avg, combine_postburn$ndvi_postburn_avg)

#snow effect average for burned 
ndsi_burned_avg <- full_wide[c("DateTime", "ndsi", "month", "year", "site")] %>%
  filter(month %in% var.snow_months) %>%
  filter(site %in% "burned") %>%
  group_by(site, year) %>%
  summarize(ndsi_burned_avg=mean(ndsi))
#snow effect average for unburned
ndsi_unburned_avg <- full_wide[c("DateTime", "ndsi", "month", "year", "site")] %>%
  filter(month %in% var.snow_months) %>%
  filter(site %in% "unburned") %>%
  group_by(site, year) %>%
  summarize(ndsi_unburned_avg=mean(ndsi))

#vegetation effect average for burned 
ndvi_burned_avg <- full_wide[c("DateTime", "ndvi", "month", "year", "site")] %>%
  filter(month %in% var.growth_months) %>%
  filter(site %in% "burned") %>%
  group_by(site, year) %>%
  summarize(ndvi_burned_avg=mean(ndvi))

#vegetation effect average for unburned
ndvi_unburned_avg <- full_wide[c("DateTime", "ndvi", "month", "year", "site")] %>%
  filter(month %in% var.growth_months) %>%
  filter(site %in% "unburned") %>%
  group_by(site, year) %>%
  summarize(ndvi_unburned_avg=mean(ndvi))

#combining data for burned
combined_burned <- inner_join(ndvi_burned_avg, ndsi_burned_avg)

#combining data for unburned
combined_unburned <- inner_join(ndvi_unburned_avg, ndsi_unburned_avg)

#correlation for burned data
cor(combined_burned$ndvi_burned_avg, combined_burned$ndsi_burned_avg)

#correlation for unburned data
cor(combined_unburned$ndvi_unburned_avg, combined_unburned$ndsi_unburned_avg)

#graphing the data 
Preburngraph <- ggplot(combine_preburn, aes(x=ndsi_preburn_avg, y=ndvi_preburn_avg)) +
  geom_point() + 
  theme_few() + 
  scale_color_few() 

Postburngraph <- ggplot(combine_postburn, aes(x=ndsi_postburn_avg, y=ndvi_postburn_avg)) +
  geom_point() + 
  theme_few() + 
  scale_color_few() 

Burnedgraph <- ggplot(combined_burned, aes(x=ndsi_burned_avg, y=ndvi_burned_avg)) +
  geom_point() + 
  theme_few() + 
  scale_color_few() 

Unburnedgraph <- ggplot(combined_unburned, aes(x=ndsi_unburned_avg, y=ndvi_unburned_avg))+
  geom_point() + 
  theme_few() + 
  scale_color_few() 

#Plotting the data in one frame
Plot <- ggarrange(Preburngraph, Postburngraph, Burnedgraph, Unburnedgraph,
                 labels = c("Pre-burned", "Post-burned", "Burned", "Unburned"),
                 ncol = 2, nrow = 2)

Plot
                  
```

For pre-burn snow effect, the correlation is 0.091.

For post-burn snow effect, the correlation is 0.244.

For un-burned snow effect, the correlation is -0.031

For burned snow effect, the correlation is 0.087.

The snow effect from question 2 is a more generalized correlation. While all of these variables have a weak correlation, this question isolates the different variables in the data to help determine the relationship between snow coverage and vegetation growth. The post-burn condition has the strongest correlation between the previous year's snow coverage and the vegetation growth for the following year. The weakest correlation between snow coverage and vegetation growth was for the un-burned condition. This is interesting because this could potentially mean that the Hayman Fire has pushed the vegetation growth to rely more on snow coverage to get their nutrients. This is still a weak correlation, so more research will need to be conducted to determine if this is true or not; there are many other variables in the environment that could have caused this correlation.
 

## Question 4

What month is the greenest month on average? 


```{r warning=F, message=F}
ndvigreenest <- full_wide[c("DateTime", "ndvi", "month", "year", "site")] %>%
  group_by(month) %>%
  summarize(ndvi=mean(ndvi))

ggplot(ndvigreenest, aes(x=month, y=ndvi)) +
  geom_bar(stat="identity", fill="#009E73") +
  theme_minimal()
  
```

The greenest month on average is August.

## Question 5) 

What month is the snowiest on average?

```{r warning=F, message=F}
ndsisnowiest <- full_wide[c("DateTime", "ndsi", "month", "year", "site")] %>%
  group_by(month) %>%
  summarize(ndsi=mean(ndsi))

ggplot(ndsisnowiest, aes(x=month, y=ndsi)) +
  geom_bar(stat="identity", fill="#56B4E9") +
  theme_minimal() 
```

The snowiest month on average is January. 