# Snow Data Example

In this assignment, I explored web scraping, different functions and iterations by using a data set from the Center for Snow and Avalanche Studies  [Website](https://snowstudies.org/archived-data/) and read a table in. This table contains links to data I want and to programatically download for three sites. I don't know much about these sites, but they contain incredibly rich snow, temperature, and precipitation data. 

```{r snowsetup, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)

```

## Reading an html 

### Extract CSV links from webpage

```{r snowfunctionsetup}

site_url <- 'https://snowstudies.org/archived-data/'

#Read the web url
webpage <- read_html(site_url)

#See if we can extract tables and get the data that way
tables <- webpage %>%
  html_nodes('table') %>%
  magrittr::extract2(3) %>%
  html_table(fill = TRUE)
#That didn't work, so let's try a different approach

#Extract only weblinks and then the URLs!
links <- webpage %>%
  html_nodes('a') %>%
  .[grepl('24hr',.)] %>%
  html_attr('href')

```

## Data Download

### Download data in a for loop

```{r snowdownloadingdata}

#Grab only the name of the file by splitting out on forward slashes
splits <- str_split_fixed(links,'/',8)

#Keep only the 8th column
dataset <- splits[,8] 

#generate a file list for where the data goes
file_names <- paste0('/Users/maddiebean21/Desktop/School/ESS580A7/bookdown_final/data/03-hw-snow/SASP_24hr.csv',dataset)

for(i in 1:3){
  download.file(links[i],destfile=file_names[i])
}

downloaded <- file.exists(file_names)

evaluate <- !all(downloaded)

```


### Download data in a map

```{r snowmapping}

#Map version of the same for loop (downloading 3 files)
if(evaluate == T){
  map2(links[1:3],file_names[1:3],download.file)
}else{print('data already downloaded')}


```

## Data read-in 

### Read in just the snow data as a loop

```{r snowreadindata}
#Pattern matching to only keep certain files
snow_files <- file_names %>%
  .[!grepl('SG_24',.)] %>%
  .[!grepl('PTSP',.)]

empty_data <- list()

snow_data <- for(i in 1:length(snow_files)){
  empty_data[[i]] <- read_csv(snow_files[i]) %>%
    select(Year,DOY,Sno_Height_M)
  }

snow_data_full <- do.call('rbind',empty_data)

summary(snow_data_full)
```


### Read in the data as a map function

```{r snowmapfunction}
#making the data as a map function
our_snow_reader <- function(file){
  name = str_split_fixed(file,'/',2)[,2] %>%
    gsub('_24hr.csv','',.)
  df <- read_csv(file) %>%
    select(Year,DOY,Sno_Height_M) %>%
    mutate(site = name)
}

#creating an object with the functions
snow_data_full <- map_dfr(snow_files,our_snow_reader)

summary(snow_data_full)
```


### Plot snow data

```{r snowplotting}

#making an object for the yearly snow data points
snow_yearly <- snow_data_full %>%
  group_by(Year,site) %>%
  summarize(mean_height = mean(Sno_Height_M,na.rm=T))

#plotting the yearly snow data
ggplot(snow_yearly,aes(x=Year,y=mean_height,color=site)) + 
  geom_point() +
  ggthemes::theme_few() + 
  ggthemes::scale_color_few()+
  labs(x="Mean Height", y="Year", title="Yearly Snow Data")
```

## Extracting the meteorological data URLs

I used the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets. 

```{r snowextracting}
#creating values for the meteorological data URLs
links_hw <- webpage %>%
  html_nodes('a') %>%
  .[grepl('forcing',.)] %>%
  html_attr('href')
links_hw
```


## Download the meteorological data. 

Next, I used the `download_file` and `str_split_fixed` commands, along with a for loop, to download the data and saved it in our data folder.

```{r, snowdownloader}

# grab only the name of the file by splitting out on forward slashes
splits_hw <- str_split_fixed(links_hw, '/', 8)

forcingdataset <- splits_hw[,8] %>% 
  gsub('.txt','',.)

file_names_hw <- paste0('/Users/maddiebean21/Desktop/School/ESS580A7/bookdown_final/data/03-hw-snow/SBB_SASP_Forcing_Data', forcingdataset)

# creating a for loop
for(i in 1:length(file_names_hw)){
  download.file(links_hw[i], destfile=file_names_hw[i])
}

file_names_hw

```

## Custom function writing

I wrote a custom function in order to read in the data and append a site column to the data. 

```{r snowfunction}

# this code grabs the variable names from the metadata pdf file
library(pdftools)
q3_headers <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:26] %>%
  str_trim(side = "left")
q3_headers

#creating a function
q3_reader <- function(file){
  name = str_split_fixed(file, '/', 2)[,2] 
  name2 = str_split_fixed(file, '_', 4)[,2]
  q3test = read.delim(file, header = FALSE, sep ="", col.names = q3_headers, skip = 4) %>%
  select(1:14) %>%
  mutate(site=name2)
}

```

## Summary of meteorlogical files

I used the `map` function to read in both meteorological files and then displayed a summary of my tibble.

```{r snowmetfiles, warning = FALSE}

#reading in the forcing data 
forcing_data_full <- map_dfr(file_names_hw, q3_reader)

summary(forcing_data_full)
```


##Average yearly temperature 

I made a line plot of mean temp by year by site (using the `air temp [K]` variable). What is suspicious about this data is temperature variation at the beginning of the graph, especially since the temperature is in Kelvin. This eludes that temperature was most likely not collected in the earlier years. 

```{r snowlineplot, warning = FALSE}
#creating an object to plot, grabbing the mean temperature by year by site
q5_yearly <- forcing_data_full %>%
  group_by(year,site) %>%
  summarize(mean_temp_k = mean(air.temp..K.,na.rm=T))
 
#plotting the mean temperature by year by site
ggplot(q5_yearly) + 
  geom_line(aes(x=year,y=mean_temp_k,color=site)) + 
  ggthemes::theme_few() + 
  ggthemes::scale_color_few()
```

What is suspicious about this data is temperature variation at the beginning of the graph, especially since the temperature is in Kelvin. This eludes that temperature was most likely not collected in the earlier years.

## Monthy average temperature plot

Here, i wrote a function in order to make line plots of monthly average temperature at each site for a given year. I used a for loop to make these plots for 2005 to 2010. Both sites follow the same trend for each month, however, the SBSP site is never warmer than SASP site. There are times where are almost the same temperature, but SBSP never exceeds SASP.

```{r snowfunctionplot}

#creating a function for the monthly average temperature
forcingmonthyear <- function(forcing_data_full, year){
  monthlytemp<-forcing_data_full %>%
    group_by(month,year,site) %>%
    summarize(monthly_air_temp = mean(air.temp..K.,na.rm=T)) %>%
    filter(yr == year)
  
  #plotting the function
  plots <- ggplot(monthlytemp, aes(x = month, y = monthly_air_temp, color = site)) +
    geom_line(size=2)+
    facet_wrap(~year)
    labs(title= monthlytemp$year,
         x = "Month",
         y = "Temperature (K)")
  
  print(plots)
}

years <- c(2005,2006,2007,2008,2009,2010)
  
#creating a for loop
for (yr in years) {
  forcingmonthyear(forcing_data_full, year)
}
  
```

## Average daily precipitation

For this graph, I grouped the data by day by site to get the daily temperature. 

```{r snowbonus}

# making an object for daily temperature
dailytemp<-forcing_data_full %>%
    group_by(day,site) %>%
    summarize(daily_air_temp = mean(air.temp..K.,na.rm=T))

#plotting bonus
ggplot(dailytemp, aes(x=day, y=daily_air_temp, color=site))+
  geom_line()+
  labs(x='Day of the Month', y='Daily Air Temperature', title = "Daily Temperature")
```


